Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-03-30 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@488 3fc257bb-8550-0410-81b3-a5c103da83f7

@@

@@

Discussion of triangle finding algorithm.

@@

 \caption{MapReduce algorithm for triangle finding.}

>>>>

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> 1 Map: \> Generate $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> 1 Collate \\
\> 1 Reduce: \> Remove duplicate edges \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

@@

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

@@

 \label{fig:pr}
\end{figure}

The MapReduce implementation performs two {\it map()} operations to
initialize the graph matrix $A$ and PageRank vector $x$.  A {\it
collate()} operation gathers all row entries $a_{ij}$ with their
associated $x_i$ entry, and a {\it reduce()} computes $a_{ij} x_i$.  A
second {\it collate()} gathers, for each $j$, all contributions to the
column sum $\sum a_{ij} x_i$, which is computed by a second {\it
reduce()}.  MPI\_Allreduce calls are used to compute global norms and
residuals.

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the three edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all such
triplets, assuming an input graph of undirected edge keys $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
generated matrix.  This exposition follows the triangle-finding
algorithm presented in \cite{Cohen}.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxxxxxxxxxx\=xxx\=xxx\=xxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

1 Copy: \> $G_0$ = copy of edge KV pairs from input graph \\
1 Map: \> Convert edges to vertices \\
           \> \> input Key = $(V_i,V_j)$, Value = NULL \\
           \> \> output Key = $V_i$, Value = $V_j$ \\
           \> \> output Key = $V_j$, Value = $V_i$ \\
1 Collate \\
1 Reduce: \> Add first degree to one vertex in edge \\
              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> for each V in MultiValue: \\
              \> \> \> if $V_i < V$: output Key = $(V_i,V)$, Value = $(D_i,0)$ \\
              \> \> \> else: output Key = $(V,V_i)$, Value = $(0,D_i)$ \\
2 Collate \\
2 Reduce: \> Add second degree to other vertex in edge \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
              \> \> output Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
3 Map: \> Low degree vertex emits edges \\
           \> \> if $D_i < D_j$: output Key = $V_i$, Value = $V_j$ \\
           \> \> else if $D_j < D_i$: output Key = $V_j$, Value = $V_i$ \\
           \> \> else: output Key = $V_i$, Value = $V_j$ \\
3 Collate \\
3 Reduce: \> Emit angles of each vertex \\
              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> for each $V_1$ in MultiValue: \\
	      \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
	      \> \> \> \> if $V_1 < V_2$: output Key = $(V_1,V_2)$, Value = $V_i$ \\
	      \> \> \> \> else: output Key = $(V_2,V_1)$, Value = $V_i$ \\
4 Add: \> Add $G_0$ edge KV pairs to angle KV pairs \\
4 Collate \\
4 Reduce: \> Emit triangles \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
              \> \> if NULL exists in MultiValue: \\
	      \> \> \> for each non-NULL V in MultiValue: \\
	      \> \> \> \> output Key = $(V_i,V_j,V)$, Value = NULL

@@

 \caption{MapReduce algorithm for triangle enumeration.}

 \label{fig:tri}
\end{figure}

The initial step is to store a copy of the graph edges as key/value
(KV) pairs in a auxiliary MapReduce object $G_0$, for use later in the
algorithm.  The first {\it map()} operation converts edge keys to
vertex keys with edge values.  After the {\it collate()}, each vertex
has a list of vertices it is connected to; the first {\it reduce()}
can thus flag one vertex $V_i$ in each edge with a degree count $D_i$.
The second {\it collate()} and {\it reduce()} assign a degree count
$D_j$ to the other vertex in each edge.  In the third {\it map()},
only the lower-degree vertex in each edge emits its edges as key/value
(KV) pairs.  The task of the third {\it reduce()} is to emit
``angles'' for each of these low-degree vertices.  An ``angle'' is a
root vertex $V_i$, with two edges to vertices $V_1$ and $V_2$, i.e. a
triangle without the third edge $(V_1,V_2)$.  The {\it reduce()} emits
a list of all angles of vertex $V_i$, by a double loop over the edges
of $V_i$.  Note that the aggregate volume of KV pairs emitted at this
stage is minimized by having only the low-degree vertex in each edge
generate angles.

In stage 4, the KV pairs in the original graph $G_0$ are added to the
current working set of KV pairs.  The KV pairs in $G_0$ are edges that
complete triangles for the angle KV pairs just generated.  After the
fourth {\it collate()}, a pair of vertices $(V_i,V_j)$ is the key, and
the multivalue is the list of all root vertices in angles that contain
$V_i$ and $V_j$.  If the multivalue also contains a NULL, contributed
by $G_0$, then there is a $(V_i,V_j)$ edge in the graph.  Thus all
vertices in the multivalue are roots of angles which are complete
triangles and can be emitted as a triplet key.
